---
title: "Whithin-Strain Analysis"
params:
    str: "C57"
editor: visual
toc: true
---

```{r}
#| label: setup
#| include: FALSE

# Import utility functions
source("globals.R")

# Set chunks options
knitr::opts_chunk$set(
  warning = FALSE, message = FALSE, echo = FALSE,
  fig.path = paste0("plots/", params$str),
  dev = c("png", "pdf")
)
```

## Overview

In this report we profile the meninges of `r params$str` mice in terms of Tertiary Lymphatic Structures (TLSs). The following chapters provide insight to different aspects of the TLSs, such as their location, distribution and density.

```{r}
#| label: data
figure <- readPNG("data/opac_meninges.png")
df <- read.csv("data/processed_data.csv")

df <- df %>%
  filter(Strain == params$str, Y < outlier_threshold) %>%
  mutate(Phenotype = as.factor(Phenotype),
         SexText = factor(SexText, c("Males", "Females")),
         Age = factor(Age, c("young", "middle", "old")),
         Sex.Age = factor(Sex.Age, levs))
```

```{r}
#| label: plot-size
upper_lim <- max(meninges_height, max(df$Y[df$Y < outlier_threshold]))
side_lim <- max(meninges_width, max(abs(df$X)))
lower_lim <- min(-meninges_bottom, min(df$Y))
```

## Exploratory statistics

The current analysis includes data on `r nrow(df)` lymphatic structures from `r n_distinct(df$TissueID)` mice, belonging to `r n_distinct(df$Strain)` strains, `r n_distinct(df$Age)` age categories and `r n_distinct(df$Sex)` genders, for a total of `r n_distinct(interaction(df$Strain, df$Age, df$Sex))` groups (@tbl-stats).

```{r}
#| label: tbl-stats
#| tbl-cap: Basic statistics on the TLSs from mice meninges. Size and length are measured in mm, whereas phenotype and sinus location are reported in terms of the number of structures per each phenotype or sinus location, respectively.
df %>%
  group_by(Sex, Age) %>%
  summarise(Mice = n_distinct(TissueID),
            Structures = n(),
            AvgDiameter = round(mean(Size), 1),
            AvgDistance = round(mean(Length), 1),
            `Phenotype (1-2-3)` = paste(sum(Phenotype == 1), sum(Phenotype == 2), sum(Phenotype == 3), sep = "-"),
            `Sinus (0-90-180)` = paste(sum(Sinus == 0), sum(Sinus == 90), sum(Sinus == 180), sep = "-"),
            Desc = unique(AgeText)) %>%
  knitr::kable()
```

@fig-bar shows the distribution the TLSs distribute in terms of their length and size across all tissues.

```{r}
#| label: fig-bar
#| fig-cap: Barplots of tertiary lyphatic structures distributed by (A) diameter and (B) distance from confluence in mm.
p1 <- ggplot(df, aes(x = Size)) +
  geom_boxplot() +
  labs(x = "TLS Diameter in mm", y = "TLS Count") +
  theme_classic()

p2 <- ggplot(df, aes(x = Length)) +
  geom_histogram(bins = 50) +
  labs(x = "Distance from Confluence in mm", y = "TLS Count") +
  theme_classic()

p1 / p2 +
  plot_annotation(tag_levels = "A")
```

```{r}
#| label: text-insets
text_df <- df %>%
  group_by(SexText, AgeText) %>%
  summarise(Mice = n_distinct(TissueID),
            Count = n()) %>%
  mutate(Label = paste(Count, "ELS from", Mice, "samples"))

insets <- df %>%
  split(f = .$Sex.Age, drop = TRUE) %>%
  purrr::map(~annotation_custom2(
    data = ., grob = ggplotGrob(get_inset(., my_cols)),
    xmin = -side_lim - 1, xmax = -side_lim + 3.5,
    ymin = upper_lim - 4.5, ymax = upper_lim)
)
```

## Distribution of structures

```{r}
#| label: fig-loc
#| fig-cap: Scatter plot of TLSs around meninges. Each dot represents one structure, sized by dimension, coloured by phenotype and labeled by sample. Each panel corresponds to a mice group based on sex and age category.
#| fig-width: !expr 5 + 2.75 * (n_distinct(df$Age) - 1)
#| fig-height: !expr 3.5 * n_distinct(df$Sex)
ggplot(df, aes(x = X, y = Y)) +
  lims(x = c(-side_lim, side_lim), y = c(lower_lim, upper_lim)) +
  insets +
  annotation_custom(rasterGrob(figure, 
                               width = unit(1, "npc"), 
                               height = unit(1, "npc")), 
                    -meninges_width, meninges_width,
                    -meninges_bottom, meninges_height) +
  geom_point(alpha = my_alpha, aes(colour = Phenotype, size = Size)) +
  geom_text(data = text_df, aes(label = Label),
            x = -side_lim + 2.4, y = upper_lim + 0.3, size = 3) +
  scale_colour_manual(values = my_cols) +
  scale_size_continuous(limits = size_lims, breaks = size_breaks) +
  labs(x = "Distance from Lambda (mm)", y = "Distance from Lambda (mm)") +
  facet_grid(SexText ~ AgeText) +
  theme_classic()
```

## Density of structures

```{r}
#| label: fig-dist
#| fig-cap: 2d Binned plot of tertiary limphatic structures around meninges. Colour indicates number of structures found in each hexagonal cell. Each panel corresponds to a mice group based on sex and age category.
#| fig-width: !expr 5 + 2.75 * (n_distinct(df$Age) - 1)
#| fig-height: !expr 3.5 * n_distinct(df$Sex)
ggplot(df, aes(x = X, y = Y)) +
  lims(x = c(-side_lim, side_lim), y = c(lower_lim, upper_lim)) +
  geom_bin2d(binwidth = c(bin_width, bin_width), 
             origin = c(-bin_width / 2, -bin_width / 2)) +
  annotation_custom(rasterGrob(figure, 
                               width = unit(1, "npc"), 
                               height = unit(1, "npc")), 
                    -meninges_width, meninges_width,
                    -meninges_bottom, meninges_height) +
  geom_text(data = text_df, aes(label = Label),
            x = -side_lim + 2.4, y = upper_lim + 0.3, size = 3) +
  scale_fill_gradient2(low = col_low, high = col_high, limits = col_lims) +
  labs(x = "X in mm", y = "Y in mm") +
  facet_grid(SexText ~ AgeText) +
  theme_classic()
```

## Distribution and Density of structures

```{r}
#| label: fig-loc-dens
#| fig-cap: Density plot of TLSs around meninges. Each dot represents one structure, sized by dimension, coloured by phenotype. Colour indicates the fraction of structures found in each pixel. Each panel corresponds to a mice group based on sex and age category.
#| fig-width: !expr 5 + 2.75 * (n_distinct(df$Age) - 1)
#| fig-height: !expr 3.5 * n_distinct(df$Sex)
ggplot(df, aes(x = X, y = Y)) +
  lims(x = c(-side_lim, side_lim), y = c(lower_lim, upper_lim)) +
  geom_hex(bins = bin_num) +
  insets +
  annotation_custom(rasterGrob(figure, 
                               width = unit(1, "npc"), 
                               height = unit(1, "npc")), 
                    -meninges_width, meninges_width,
                    -meninges_bottom, meninges_height) +
  geom_point(aes(colour = Phenotype, size = Size), alpha = my_alpha) +
  geom_text(data = text_df, aes(label = Label),
            x = -side_lim + 2.4, y = upper_lim + 0.3, size = 3) +
  scale_colour_manual(values = my_cols) +
  scale_size_continuous(limits = size_lims, breaks = size_breaks) +
  scale_fill_gradient2(low = col_low, high = col_high, limits = col_lims) +
  labs(x = "Distance from Lambda (mm)", y = "Distance from Lambda (mm)", fill = "Count") +
  facet_grid(SexText ~ AgeText) +
  theme_classic()
```

## Distribution by phenotype

```{r}
#| label: fig-dist-pheno
#| fig-cap: 2d Binned plot of tertiary limphatic structures around meninges by phenotype (1, 2 or 3). Colour indicates number of structures found in each hexagonal cell. Rows correspond to a mice group based on sex and age category, whereas columns represent phenotypes.
#| fig-width: 10
#| fig-height: !expr 3 * n_distinct(df$Sex.Age)
ggplot(df, aes(x = X, y = Y)) +
  lims(x = c(-side_lim, side_lim), y = c(lower_lim, upper_lim)) +
  geom_hex(bins = bin_num) +
  annotation_custom(rasterGrob(figure, 
                               width = unit(1, "npc"), 
                               height = unit(1, "npc")), 
                    -meninges_width, meninges_width,
                    -meninges_bottom, meninges_height) +
  geom_point(aes(colour = Phenotype, size = Size), alpha = my_alpha) +
  scale_colour_manual(values = my_cols) +
  scale_size_continuous(limits = size_lims, breaks = size_breaks) +
  scale_fill_gradient2(low = col_low, high = col_high, limits = col_lims) +
  labs(x = "X in mm", y = "Y in mm", fill = "Count") +
  facet_grid(Sex.Age ~ Phenotype) +
  theme_classic()
```
